DataCamp: An Interactive Teaching Tool

### 2018-01-14 manuscript.tex
edited 'advise' section
edits to manuscript

The purpose of this article is to introduce a new interactive teaching technology, DataCamp.

\subsection{DataCamp course features}

@@

\subsubsection{Wide variety of course topics}
The DataCamp website offers free `community' courses as well as more specialized \textit{Premium} classes that require a fee. 
However, when used for academic purposes the \textit{Premium} DataCamp content is free for up to six months.
College educators can create a classroom group on the DataCamp website, add students to the group, 
assign any of the existing courses and even assign a course that they've developed themselves. 
The collection of community courses available on DataCamp include a wide range of topics from \textit{Introduction to R} to
\textit{Quantitative Risk Management in R} so novice through advanced programming students benefit.

@@

\subsubsection{Integration with Learning Management Systems}
One feature that sets DataCamp apart from competing technologies is the ability to track student progress. Educators
can set deadlines for assignments and view which sections the
students had trouble with in a detailed report. DataCamp courses can also be linked to the school's learning management system (LMS) so student scores are automatically updated to the
current grade-book, allowing for a smooth integration of student work.

\subsection{Pre-existing R courses}
In order to utilize 
DataCamp's \textit{classroom} benefits, first create an academic group on the website \url{https://www.datacamp.com/groups/education}.
After your university information has been verified and the classroom account has been approved, access is facilitated 
by either creating a group through the DataCamp website or linking directly to your LMS. You can add students to the course by their university email address on the \textit{Members} tab.
\section{Creating a DataCamp Course}
\subsection{Getting Started}
To build a course on DataCamp you will first need to create a DataCamp account by visiting \url{www.datacamp.com/teach} as well as
a GitHub account \url{https://github.com}. Every DataCamp course is linked through GitHub version control,
allowing statistics educators to easily collaborate and reproduce code. To add collaborators to the course, go to the course repository
on GitHub and under \texttt{Settings} you can enter the user names of collaborators. Collaborators who are not the primary author need 
to uncheck `My Courses' in order to view and contribute to the repository. The simplest way to begin building the course is to use the template course files from the
\texttt{Create DataCamp Course} dialog. This will automatically create a GitHub repository that is linked to DataCamp; any changes that are
made through DataCamp will update the GitHub repository and vice versa.

@@

\texttt{Readme.md} contains helpful information of how to get started, citing resources to learn more about the process.
\texttt{course.yml} is a file that contains general information about the course you're creating: title, university, description,
difficulty level, ect.
\texttt{chapter1.Rmd} is a template chapter file with examples of 'normal' iterative coding exercises and multiple choice questions.
To add more chapters to an R course, create new files in the repository and name them \texttt{chapter1.Rmd}, \texttt{chapter2.Rmd},
\texttt{chapter3.Rmd}, ect.

\subsection{Course Editing}
Educators are able to edit their courses in a number of ways but the \texttt{Teach Editor} located in the teach dashboard is the
most straight forward editor. The \texttt{Teach Editor} is specifically designed to make the creation process of a DataCamp course
as painless as possible, allowing writers to edit, preview, save changes and synchronize to a GitHub repository. Editing can
also be done through GitHub's online editor or through git locally but these methods do not allow for automatic content
viewing.
\subsection{Components of a DataCamp Course}
\subsubsection{Exercise Layout}
Figure1 is an example of an exercise in the editor and figure2 is the corresponding interface for students. Students will read the lesson information and instructions, answer the question, and submit their code; it should take approximately 5 minutes for a student to complete this exercise.
\begin{figure}
\caption{Teach Editor Interface}
\begin{Verbatim}[frame=single]
--- type:NormalExercise lang:r xp:100 skills:1 key:9a4fea2da
## Visualizing quantitative data in R

@@

\begin{figure}[h]
  \includegraphics[scale = 0.40] {preview.jpg}
\end{figure}
\subsubsection{Exercise Header}
Whenever a new exercise is added, DataCamp automatically produces the top line of the editor. The header specifies the type
of exercise (\texttt{type}), programming language (\texttt{lang}), available points (\texttt{xp}), skills learned (\texttt{skills}) and a \texttt{key}. 
The types of exercises include normal coding, video, and multiple choice exercises; we will discuss these further in the next section. In addition to R, DataCamp offers courses in other programming languages including Python and SQL for data science.
DataCamp assigns 100 points \texttt{xp} to normal coding exercises but this value can be modified depending on
difficulty level. After students have completed the exercise, they can check their personal profile to see how many points they earned.
The \texttt{skills} portion of the header refers to DataCamp's
eight different skill sets, defined by numbers 1-8. Skills and their corresponding numbers can be found in the DataCamp
documentation \url{https://www.datacamp.com/teach/documentation#tab_gamification}. 
Lastly, the \texttt{key} acts as a unique
identifier for that exercise, allowing you to modify content without losing any student data. 
In the above exercise code (refer to teach editor code), the \texttt{type} is Normal, the language is R, the available points is 100xp and the skills learned is 1 (which 
corresponds to R Programming).

\subsubsection{Exercise Body}
In the \textit{Assignment} portion of the exercise (not labeled but below the header), instructors can add additional information about the assignment that will help
students understand and complete the instructions. The \textit{instructions} section contains the actual task the student is
asked to execute; instructors can also include a \textit{hint} for extra guidance (optional). The \textit{pre-exercise code} sets up the work space for 
the student - this section contains code to import data sets, install packages and any other operation that you want to be available to
students. The \textit{sample code} section will contain code or notes that will be visible to students in the R script panel. Students also use the R script panel to write code that solves the tasks in the instructions. The correct answer to an exercise is coded in the \textit{solution} section
and when a student clicks the \texttt{Submit Code} button, their answer passes through the SCT code where their answer is checked for errors.

\subsubsection{Submission Correctness Test}
The submission correctness testing code (or SCT) automatically assesses if the correct code was submitted. If a student submits an 
incorrect answer, they will receive immediate feedback on what they are doing wrong and the system will guide them to the right answer.
DataCamp provides a helpful R package called \textit{testwhat} which acts as a guide to writing and using SCTs - available on DataCamp's GitHub 
repository: \url{https://github.com/datacamp/testwhat/wiki}. 
\textit{testwhat} contains functions that test different types of problems including loops, object definitions, printed output, functions, and other R programming problems.
The submission correctness testing code compares the ideal answer (from the solution code) to the students answer. Various
arguments can be added to the test functions to adjust the testing process and feedback.
Instructors can either use the default feedback that DataCamp automatically generates or write their own custom messages. For example, figure blank
contains the default feedback for a few possible incorrect submissions of the \texttt{hist()} function from the example exercise \textit{Visualizing quantitative data in R}.

@@

Student types:  histogram(dataset$var_name)

Default message:
Have you called hist()?

***

Student types: hist(var_name)

Default message:
Check your call of hist(). Did you correctly specify 
the argument x? Evaluating the expression you specified
caused an error.

***

Student types: hist(dataset$misspelled_var)

Default message:
Check your call of hist(). Did you correctly specify the
argument x? 
It is a NULL, while it should be a numeric vector.

\end{Verbatim}
\end{figure}
DataCamp's default feedback is meaningful and specific so a lot of the time custom 
messages are not necessary.
However, if students are struggling with a common problem then unique feedback can be very beneficial.

@@

\subsection{Advise for smoother course development}
DataCamp is fairly new and the only available information about course creation is on the DataCamp website. While writing my first DataCamp R programming course, I learned a few tricks that I wish I knew before I started.

@@

\subsubsection{Integrating R packages}
All DataCamp courses will have a course badge (image) and most will use data sets and R packages. Information on how to add data sets and images
can be found in DataCamps \textit{Teach} documentation \url{https://www.datacamp.com/teach/documentation} under the `Upload Assets' tab. However,
there is currently no information on DataCamp about integrating R packages. To use R packages in a course, you will need to create a new file
in the GitHub course repository named \texttt{requirements.r}. This file should contain a line of code with three components: \texttt{devtools::install\_version}, package name and version for 
every package you wish to include in the DataCamp course. For example, the following code will add the \texttt{ggplot2} and \texttt{mosaic} packages to the repository.

@@

\subsubsection{TeX typesetting}
LaTeX is compatible with DataCamp; course builders can use LaTeX math symbols to display Greek letters, formulas, ect. For example:
\begin{verbatim}
$H_0$: $p = 0.07$
$H_a$: $p \neq 0.07$
\end{verbatim}

>>>>

The purpose of this article is to introduce a new interactive teaching technology, DataCamp.
In addition to R, DataCamp offers courses in other programming languages including Python and SQL for data science.
\subsection{DataCamp course features}

@@

\subsubsection{Introduction to DataCamp - don't know what that note says?}
DataCamp website offers free "community" courses as well as more specialized \textit{Premium} classes that require a fee. 
However, when used for academic purposes the \textit{Premium} DataCamp content is free for up to six months.
College educators can create a classroom group on the DataCamp website, add students to the group, 
assign any of the existing courses and even assign a course that they have developed themselves. 
The collection of community courses available on DataCamp include a wide range of topics from \textit{Introduction to R} to
\textit{Quantitative Risk Management in R} so novice through advanced programming students benefit.

@@

\subsubsection{Integration with Learning Management Systems}
One feature that sets DataCamp apart from competing technologies is the ability to track student progress. Educators
can set deadlines for assignments and view which sections the
students had trouble with in a detailed report. DataCamp courses can also be linked to the school's learning management system (LMS), such as Moodle or Blackboard, so student scores are automatically updated to the
current grade-book, allowing for a smooth integration of student work.

\subsection{Pre-existing R courses}
In order to utilize 
DataCamp's \textit{classroom} benefits, first create an academic group on the website \url{https://www.datacamp.com/groups/education}.
After your university information has been verified and the classroom account has been approved, access is facilitated 
by adding students to the DataCamp group by supplying university email addresses on the \textit{Members} tab, or by creating an integration to your LMS.
\section{Creating a DataCamp Course}
\subsection{Getting Started}
To build a course on DataCamp you will first need to create a DataCamp account by visiting \url{www.datacamp.com/teach} as well as
a GitHub account (\url{https://github.com}). (Should this be in section 3.5.2? -- Every DataCamp course is linked through GitHub version control,
allowing statistics educators to easily collaborate and reproduce code.) The simplest way to begin building the course is to use the template course files from the
\texttt{Create DataCamp Course} dialog. This will automatically create a GitHub repository that is linked to DataCamp; any changes that are
made through DataCamp will update the GitHub repository and vice versa.

@@

\texttt{Readme.md} contains helpful information of how to get started, citing resources to learn more about the process.
The \texttt{course.yml} is a file that contains general information about the course you're creating: title, university, description,
difficulty level, ect.
The \texttt{chapter1.Rmd} is a template chapter file with examples of `normal' iterative coding exercises and multiple choice questions.
To add more chapters to an R course, create new files in the repository and name them \texttt{chapter1.Rmd}, \texttt{chapter2.Rmd},
\texttt{chapter3.Rmd}, ect.

\subsection{Course Editing}
Educators are able to edit their courses in a number of ways, but the \texttt{Teach Editor} located in the DataCamp dashboard is the
most straight forward editor. The \texttt{Teach Editor} is specifically designed to make the creation process of a DataCamp course
as painless as possible, allowing writers to edit, preview, save changes and synchronize to a GitHub repository. Editing can
also be done through GitHub's online editor or through git locally but these methods do not allow for automatic content
viewing.
\subsection{Components of a DataCamp Course}
\subsubsection{Exercise Layout}
Figure1 is an example of an exercise in the Teach Editor and figure2 is the corresponding interface for students. Students will read the lesson information and instructions, answer the question, and submit their code; it is recommended that each exercise takes approximately 5 minutes.
\begin{figure}
\caption{Teach Editor Interface}
\begin{Verbatim}[frame=single]
--- type:NormalExercise lang:r xp:100 skills:1 key:9a4fea2da
## Visualizing quantitative data in R

@@

\begin{figure}[h]
  \includegraphics[scale = 0.40] {preview.jpg}
\end{figure}
\subsubsection{Exercise Header}
Whenever a new exercise is added, DataCamp automatically produces the exercise header (---type:, etc.). The header specifies the type
of exercise (\texttt{type}), programming language (\texttt{lang}), available points (\texttt{xp}), skills learned (\texttt{skills}) and a \texttt{key}. 
The types of exercises include normal coding, video, and multiple choice exercises; which will be discussed further in the next section. 
DataCamp assigns 100 points \texttt{xp} by default to normal coding exercises but this value can be modified depending on
difficulty level. After students have completed the exercise, they can check their personal profile to see how many points they earned.
The \texttt{skills} portion of the header refers to DataCamp's
eight different skill sets, defined by numbers 1-8. Skills and their corresponding numbers can be found in the DataCamp
documentation \url{https://www.datacamp.com/teach/documentation#tab_gamification}. 
Lastly, the \texttt{key} acts as a unique
identifier for that exercise, allowing you to modify content without losing any student data. 
In Figure 1, the \texttt{type} is Normal, the language is R, the available points is 100xp and the skills learned is 1 (which 
corresponds to R Programming).

\subsubsection{Exercise Body}
In the \textit{Assignment} portion of the exercise (not labeled but below the header), instructors can add additional information about the assignment that will help
students understand and complete the instructions, which is displayed in the top left panel of Figure 2. The \textit{instructions} section contains the actual task the student is
asked to execute; instructors can also include a \textit{hint} for extra guidance (bottom left panel Figure 2). The \textit{pre-exercise code} sets up the work space for 
the student - this section contains code to import data sets, install packages and any other operation that you want to be available but not visible to
students. The \textit{sample code} section will contain code or notes that will be visible to students in the R script panel (top right figure 2). Students also use the R script panel to write code that solves the tasks in the instructions. The correct answer to an exercise is coded in the \textit{solution} section
and when a student clicks the \texttt{Submit Code} button, their answer passes through the R console, and then the Submission Correctness Testing (SCT) code where their answer is checked for errors.

\subsubsection{Submission Correctness Test}
The submission correctness testing code (or SCT) automatically assesses if the correct code was submitted. If a student submits an 
incorrect answer, they will receive immediate feedback on what they are doing wrong and the system will guide them to the right answer.
DataCamp provides a helpful R package called \textit{testwhat} which acts as a guide to writing and using SCTs - available on DataCamp's GitHub 
repository: \url{https://github.com/datacamp/testwhat/wiki}. 
The \textit{testwhat} package contains functions that test different types of problems including loops, object definitions, printed output, functions, and other R programming problems.
The submission correctness testing code compares the ideal answer (from the solution code) to the students answer. Various
arguments can be added to the test functions to adjust the testing process and feedback.
Instructors can either use the default feedback that DataCamp automatically generates or write their own custom messages. For example, figure blank
contains the default feedback for a few possible incorrect submissions of the \texttt{hist()} function. 
Suppose the data set is called \texttt{birthdata} and the variable of interest is \texttt{weight}.

@@

It's weird but I get errors when I name the data set birth_data.

Student types:                  Default message:

histogram(birthdata$weight)    Have you called hist()?

hist(weight)                    Check your call of hist(). Did you 
                                correctly specify the argument x? 
                                Evaluating the expression you specified
                                caused an error.

hist(birthdata$wieght)         Check your call of hist(). Did you
                                correctly specify the argument x? It is
                                a NULL, while it should be a numeric vector.

\end{Verbatim}
\end{figure}

\textit{I might want to add one where student types something with the wrong symbol but I can't get onto the datacamp course for some reason.}

DataCamp's default feedback is meaningful and specific so a lot of the time custom 
messages are not necessary.
However, if students are struggling with a common problem then unique feedback can be very beneficial.

@@

\begin{verbatim}
Multiple choice exercises use the SCT function called  \texttt{test_mc} 
which requires two arguments: \texttt{correct} and \texttt{feedback_msgs}. 
The \texttt{correct} argument specifies the position of the correct answer;
if the correct answer is the second option listed, then set \texttt{correct = 2}. 
The \texttt{feedback_msgs} argument is a vector of strings that contain feedback 
messages corresponding to the positions of the listed options.

\end{verbatim}
\subsection{Advise for smoother course development}
DataCamp is fairly new and the only available information about course creation is on the DataCamp website. While the documentation on DataCamp is comprehensive, here are some tips for developing a first course.

@@

\subsubsection{Collaboration for course development}
To add collaborators to the course, go to the course repository
on GitHub and under \texttt{Settings} you can enter the user names of collaborators. Collaborators who are not the primary author need 
to uncheck `My Courses' in order to view and contribute to the repository.
\subsubsection{Integrating R packages}
Information on how to add data sets and images to a course
can be found in DataCamps \textit{Teach} documentation \url{https://www.datacamp.com/teach/documentation} under the `Upload Assets' tab. However,
there is currently no information on DataCamp about integrating R packages. To use R packages in a course, you will need to create a new file
in the GitHub course repository named \texttt{requirements.r}. This file should contain a line of code with three components: \texttt{devtools::install\_version}, package name and version for 
every package you wish to include in the DataCamp course. For example, the following code will add the \texttt{ggplot2} and \texttt{mosaic} packages to the repository.

@@

\subsubsection{TeX typesetting or Markdown}
LaTeX is compatible with DataCamp; course builders can use LaTeX math symbols to display Greek letters, formulas, ect. For example:
\begin{verbatim}
$H_0$: $p = 0.07$
$H_a$: $p \neq 0.07$
\end{verbatim}
